---
title: "Exploratory Analysis"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include = FALSE}
library(tidyverse)
library(viridis)
library(plotly)
library(dplyr)

knitr::opts_chunk$set(
	message = FALSE,
	warning = FALSE,
	fig.width = 10,
  out.width = "80%"
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

Our dataset does not contain data for October, November and December of the year 2022. We are interested in the following main felonies: Burglary, Felony Assault, Grand Larceny, Grand Larceny of Motor Vehicle, Murder and Non-negligent Manslaughter, Rape, and Robbery.

```{r read&tidy, include = FALSE}
complaint =
  tibble(
    files = list.files("./Data_preprocessing/data/"),
    path = str_c("./Data_preprocessing/data/", files)
  ) %>% 
  mutate(data = map(path, read_csv)) %>% 
  unnest(data) %>% 
  select(-files, -path, -cmplnt_num, -ky_cd, -atpt_cptd, -jurisdiction, -jurisdiction_code) %>% 
  mutate(day_of_week = strftime(as.Date(date, format = "%m/%d/%Y"),"%a")) %>% 
  separate(date, into = c("month", "day", "year"), sep = "/", convert = TRUE) %>% 
  mutate(
    month = month.abb[month],
    hour = substr(time, start = 1, stop = 2),
    borough = ifelse(borough == "(null)", NA, borough),
    susp_sex = ifelse(susp_sex %in% c("M", "F"), susp_sex, NA),
    susp_age_group = ifelse(susp_age_group %in% c("<18", "18-24", "25-44", "45-64", "65+"), susp_age_group, NA),
    susp_race = ifelse(susp_race %in% c("UNKNOWN", "(null)"), NA, susp_race),
    vic_sex = ifelse(vic_sex %in% c("M", "F"), vic_sex, NA),
    vic_age_group = ifelse(vic_age_group %in% c("<18", "18-24", "25-44", "45-64", "65+"), vic_age_group, NA),
    vic_race = ifelse(vic_race %in% c("UNKNOWN", "(null)"), NA, vic_race),
    offense = ifelse(offense %in% c("BURGLARY", "FELONY ASSAULT", "GRAND LARCENY", "GRAND LARCENY OF MOTOR VEHICLE", "MURDER & NON-NEGL. MANSLAUGHTER", "RAPE", "ROBBERY"), offense, NA)
  ) %>% 
  select(-day, -time) %>% 
  mutate(
    year = as.factor(year),
    month = factor(month, levels = month.abb),
    day_of_week = factor(day_of_week, levels = c("Sun", "Mon", "Tue", "Wed", "Thu", "Fri", "Sat")),
    hour = factor(hour),
    precinct = as.factor(precinct),
    borough = as.factor(borough),
    level = as.factor(level),
    offense = as.factor(offense),
    susp_sex = as.factor(susp_sex),
    susp_age_group = as.factor(susp_age_group),
    susp_race = as.factor(susp_race),
    vic_sex = as.factor(vic_sex),
    vic_age_group = as.factor(vic_age_group),
    vic_race = as.factor(vic_race),
  )
```

### Felony Frequency by Borough

```{r plot_borough}
complaint %>% 
  drop_na(borough, offense) %>% 
  group_by(borough) %>% 
  summarize(n_obs = n()) %>% 
  ggplot(aes(x = reorder(borough, -n_obs), y = n_obs, fill = reorder(borough, -n_obs))) +
  geom_bar(stat = 'identity') +
  labs(
    title = "Frequency of Felonies by Borough (2016-2022)",
    x = "Borough",
    y = "Frequency"
  ) +
  theme(legend.position = "none")
```

### Felony Frequency by Crime Type

```{r plot_offense}
complaint %>% 
  drop_na(offense) %>% 
  group_by(offense) %>% 
  summarize(n_obs = n()) %>% 
  mutate(offense = reorder(offense, n_obs)) %>% 
  ggplot(aes(x = offense, y = n_obs, fill = offense)) +
  geom_bar(stat = 'identity') +
  labs(
    title = "Frequency of Felonies (2016-2022)",
    x = "Felony",
    y = "Frequency"
  ) +
  coord_flip() +
  theme(legend.position = "none")
```

### Trend of Felony Frequency

```{r plot_trend}
complaint %>% 
  filter(year != 2022) %>% 
  drop_na(offense) %>% 
  group_by(offense, year) %>% 
  summarize(n_obs = n()) %>% 
  ggplot(aes(x = year, y = n_obs, fill = offense)) +
  geom_bar(stat = 'identity') +
  labs(
    title = "Frequency of 7 Felonies by Year (2016-2021)",
    x = "Year",
    y = "Frequency",
    fill = "Felony"
  ) +
  theme(legend.position = "bottom") +
  guides(fill = guide_legend(nrow = 4,byrow = TRUE))
```

### Trend of Proportion of Felony Frequency by Demographics of Victim

#### Sex

```{r plot_vic_sex}
complaint %>% 
  drop_na(offense, vic_sex) %>% 
  group_by(year, vic_sex) %>% 
  summarize(n_obs = n()) %>% 
  group_by(year) %>% 
  summarize(vic_sex, percentage = n_obs/sum(n_obs)) %>% 
  ggplot(aes(x = year, y = percentage, fill = vic_sex)) +
  geom_bar(stat = 'identity') +
  labs(
    x = "Year",
    y = "Proportion",
    title = "Proportions of Felonies by Victim Sex and Year",
    fill = "Victim Sex"
  )
```

#### Age Group

```{r plot_vic_age_group}
complaint %>% 
  drop_na(offense, vic_age_group) %>% 
  group_by(year, vic_age_group) %>% 
  summarize(n_obs = n()) %>% 
  group_by(year) %>% 
  summarize(vic_age_group, percentage = n_obs/sum(n_obs)) %>% 
  ggplot(aes(x = year, y = percentage, fill = vic_age_group)) +
  geom_bar(stat = 'identity') +
  labs(
    x = "Year",
    y = "Proportion",
    title = "Proportions of Felonies by Victim Age Group and Year",
    fill = "Victim Age Group"
  )
```

#### Race

```{r plot_vic_race}
complaint %>% 
  drop_na(offense, vic_race) %>% 
  group_by(year, vic_race) %>% 
  summarize(n_obs = n()) %>% 
  group_by(year) %>% 
  summarize(vic_race, percentage = n_obs/sum(n_obs)) %>% 
  ggplot(aes(x = year, y = percentage, fill = vic_race)) +
  geom_bar(stat = 'identity') +
  labs(
    x = "Year",
    y = "Proportion",
    title = "Proportions of Felonies by Victim Race and Year",
    fill = "Victim Race"
  )
```

### Month, Day of the Week, Time

#### Average Daily Felony Frequency by Year and Month

```{r plot_year&month}
num_days = function(month, year) {
  
  year = as.integer(year)
  months = 1:12
  names(months) = month.abb
  month = months[month]
  
  as.numeric(strftime(as.Date(paste(year + month %/% 12, month %% 12 + 1, "01", sep = "-")) - 1, "%d"))
  
}

complaint %>% 
  drop_na(offense) %>% 
  mutate(year = fct_rev(year)) %>% 
  group_by(year, month) %>% 
  summarize(mean_freq = n() / num_days(month, year)) %>% 
  plot_ly(
    x = ~month, y = ~year, z = ~mean_freq,
    type = "heatmap"
  ) %>% 
  colorbar(x = 1, y = 1) %>% 
  layout(
    title = "Average Daily Felony Frequency by Year and Month",
    xaxis = list(title = "Month"),
    yaxis = list(title = "Year")
  )
```

#### Average Hourly Felony Frequency by Time of the Week

```{r plot_week&time}
complaint %>% 
  drop_na(offense) %>% 
  mutate(
    day_of_week = fct_rev(day_of_week)
  ) %>% 
  group_by(hour, day_of_week) %>% 
  summarize(mean_freq = n() / 352) %>% 
  plot_ly(
    x = ~hour, y = ~day_of_week, z = ~mean_freq,
    type = "heatmap"
  ) %>% 
  colorbar(x = 1, y = 1) %>% 
  layout(
    title = "Average Hourly Felony Frequency by Time of the Week",
    xaxis = list(title = "Hour of the Day"),
    yaxis = list(title = "Day of the Week")
  )
```

