---
title: "Statistical Testing"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

\ \par

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = FALSE, message = FALSE)

library(tidyverse)
library(purrr)
library(modelr)
library(dbplyr)
library(plyr)
library(lubridate)
```

\ \par

```{r include=FALSE}
# Dataset preprocessing
data_source <- list.files("./Data")     
data_source 
dir <- paste("./Data/", data_source, sep = "")                 
n <- length(dir) 

merge_data <- read.csv(file = dir[1],header = T,sep = ",")   
merge_data <- cbind(dir[1], merge_data )
merge_data <- rename(merge_data, c("dir[1]" = "yyyymm"))

for (i in 2:7){
  new.data <- read.csv(file = dir[i], header = T, sep = ",")
  new.data <- cbind(dir[i], new.data)
  new.data <- rename(new.data, c("dir[i]" = "yyyymm"))
  merge_data <- rbind(merge_data,new.data)
}
```

```{r include=FALSE}
# Data cleaning, drop data with unknown.
# Filter data with unknown.
# We assume the world is in a covid state when year >= 2020.
tidydata1 = merge_data %>%
  na.omit() %>%
  filter(!susp_sex == "U") %>%
  filter(!susp_race == "UNKNOWN") %>%
  filter(!susp_age_group == "UNKNOWN") %>%
  filter(!vic_race == "UNKNOWN") %>%
  filter(!vic_age_group == "UNKNOWN") %>%
  separate(date,into = c("mon","day","year"),sep = "/") %>%
  mutate(jurisdiction_code=as.numeric(jurisdiction_code)) %>%
  mutate(
    susp_sex = ifelse(susp_sex == "M",1,0),
    vic_sex = ifelse(vic_sex == "M",1,0),
    jurisdiction_code = ifelse(jurisdiction_code >= 3,3,jurisdiction_code + 0)
    ) %>%
  mutate(jurisdiction_code = as.factor(jurisdiction_code)) %>%
  mutate(covid_state = ifelse(year >= 2020,1,0))

```

```{r include=FALSE}
tidydata3 = tidydata1 %>%
  mutate(
   time = substring(time, 1,2)
  )

tidydata3 = tidydata3 %>% 
 mutate(
   time1 = substring(time,1,1),
   time2 = substring(time,2,2),
   time3 = ifelse(time1 == "0","",time1),
   time4 = paste(time3,time2,sep = ""),
   time5 = as.numeric(time4)
 ) 

tidydata3 = tidydata3 %>% 
  mutate(time_range = case_when(
      (time5 > 22) | (time5 <= 6) ~ 0,
      (time5 > 6) & (time5 <= 14) ~ 1,
      (time5 > 14) & (time5 <= 22) ~ 2
    ))
```


In this section we use statistical testing methods to conduct in-depth analysis towards the distribution and different features of the NYC crime data.


## When dangers define as the criminal events happen by day

First, we define the "danger" as the _total number of crime event_ occur in _one day_

### 1.Does dangers correlate to the criminal's gender?
```{r}
crimebyday_men = tidydata1 %>% 
  group_by(year,mon,day) %>% 
  filter(susp_sex == 1) %>% 
  dplyr::summarize(
    n_bydate = n(),
  ) %>% 
  unique()

crimebyday_women = tidydata1 %>% 
  group_by(year,mon,day) %>% 
  filter(susp_sex == 0) %>% 
  dplyr::summarize(
    n_bydate = n(),
  ) %>% 
  unique()

t.test(crimebyday_men$n_bydate,crimebyday_women$n_bydate)
```

The Welch two sample t-test shows that there is a significant difference between criminal genders on their daily criminal events, with male criminals much more than female.

### 2.Does dangers correlate to the the victim's gender?
```{r}
crimebyday_men = tidydata1 %>% 
  group_by(year,mon,day) %>% 
  filter(vic_sex == 1) %>% 
  dplyr::summarize(
    n_bydate = n(),
  ) %>% 
  unique()

crimebyday_women = tidydata1 %>% 
  group_by(year,mon,day) %>% 
  filter(vic_sex == 0) %>% 
  dplyr::summarize(
    n_bydate = n(),
  ) %>% 
  unique()

t.test(crimebyday_men$n_bydate,crimebyday_women$n_bydate)
```

The Welch two sample t-test also shows significant difference between victim's gender. Female victims are significantly higher than male on daily criminal events.

### 3.Does dangers correlate to the criminals' race?
```{r}
crimebyday_black = tidydata1 %>% 
  group_by(year,mon,day) %>% 
  filter(susp_race == "BLACK") %>% 
  dplyr::summarize(
    n_bydate = n(),
  ) %>% 
  unique()

crimebyday_white = tidydata1 %>% 
  group_by(year,mon,day) %>% 
  filter(susp_race == "WHITE") %>% 
  dplyr::summarize(
    n_bydate = n(),
  ) %>% 
  unique()
t.test(crimebyday_black$n_bydate,crimebyday_white$n_bydate)
```

The Welch two sample t-test also shows significant difference between victim's gender. White criminals are significantly lower than black criminals on daily criminal events.

### 4.What about other criminals' race?
```{r}
crimebyday_aspa = tidydata1 %>% 
  group_by(year,mon,day) %>% 
  filter(susp_race == "ASIAN / PACIFIC ISLANDER") %>% 
  dplyr::summarize(
    n_bydate = n(),
  ) %>% 
  unique()

crimebyday_white = tidydata1 %>% 
  group_by(year,mon,day) %>% 
  filter(susp_race == "WHITE") %>% 
  dplyr::summarize(
    n_bydate = n(),
  ) %>% 
  unique()

t.test(crimebyday_aspa$n_bydate,crimebyday_white$n_bydate)
```

The test shows that the number of ASIAN / PACIFIC ISLANDER race of criminals are less than white criminals.


### 5.Crime events number by covid status
```{r}
crimebyday_covid0 = tidydata1 %>% 
  group_by(year,mon,day) %>% 
  dplyr::summarize(
    n_bydate = n(),
    covid_state = covid_state
  ) %>% 
  unique()

casenum_byday_0 = ifelse(crimebyday_covid0$covid_state == 0, crimebyday_covid0$n_bydate,NA) %>% 
  na.omit()
casenum_byday_1 = ifelse(crimebyday_covid0$covid_state == 0, NA, crimebyday_covid0$n_bydate) %>% 
  na.omit()

t.test(casenum_byday_0,casenum_byday_1)
```

The difference is significant. Among the overall status of covid we can see there is a downtrend before covid and after covid.

### 6.Compare crime events number by time range
```{r}
crimebyday_time0 = tidydata3 %>%
  filter(time_range == 0) %>% 
  group_by(year,mon,day) %>% 
  dplyr::summarize(
    n_bydate = n()
  ) %>% 
  unique()

crimebyday_time1 = tidydata3 %>%
  filter(time_range == 1) %>% 
  group_by(year,mon,day) %>% 
  dplyr::summarize(
    n_bydate = n()
  ) %>% 
  unique()

crimebyday_time2 = tidydata3 %>%
  filter(time_range == 2) %>% 
  group_by(year,mon,day) %>% 
  dplyr::summarize(
    n_bydate = n()
  ) %>% 
  unique()


t.test(crimebyday_time0$n_bydate,crimebyday_time1$n_bydate)
t.test(crimebyday_time0$n_bydate,crimebyday_time2$n_bydate)
t.test(crimebyday_time1$n_bydate,crimebyday_time2$n_bydate)
```

From the test, we can see a significant different number of criminal events occur at different time range:22-6 6-14 14-22 during a day. And the relation can be conclude as:  
22-6 < 6-14 < 14-22

## Dangers define as the FELONY events happen by day

Second, we redefine the "danger" as the _total number of felony crime event_ occur in _one day_

### 1. About races
```{r}
crimebyday_aspa = tidydata1 %>% 
  filter(level == "FELONY") %>%
  group_by(year,mon,day) %>% 
  filter(susp_race == "ASIAN / PACIFIC ISLANDER") %>% 
  dplyr::summarize(
    n_bydate = n(),
  ) %>% 
  unique()

crimebyday_white = tidydata1 %>% 
  filter(level == "FELONY") %>%
  group_by(year,mon,day) %>% 
  filter(susp_race == "WHITE") %>% 
  dplyr::summarize(
    n_bydate = n(),
  ) %>% 
  unique()

t.test(crimebyday_aspa$n_bydate,crimebyday_white$n_bydate)
```

The races is follow the same as normal event assumption.

### 2. About covid state
```{r}
compare_date_data3 = tidydata3 %>% 
  filter(level == "FELONY") %>% 
  group_by(year,mon,day) %>% 
  dplyr::summarize(
    n_bydate = n(),
    covid_state = covid_state
  ) %>% 
  unique()

casenum_byday_0 = ifelse(compare_date_data3$covid_state == 0,compare_date_data3$n_bydate,NA) %>% 
  na.omit()
casenum_byday_1 = ifelse(compare_date_data3$covid_state == 0,NA,compare_date_data3$n_bydate) %>% 
  na.omit()

t.test(casenum_byday_0,casenum_byday_1)
```

Based on testing results, we can see after covid, the probability of happen FELONY is greater than the time before covid.

### 3. About time range
```{r}
compare_date_data3 = tidydata3 %>% 
  filter(level == "FELONY") %>% 
  group_by(year,mon,day,time_range) %>% 
  dplyr::summarize(
    n_bydate = n(),
    covid_state = covid_state
  ) %>% 
  unique()

casenum_byday_0=ifelse(compare_date_data3$time_range==0,compare_date_data3$n_bydate,NA) %>% 
  na.omit()
casenum_byday_1=ifelse(compare_date_data3$time_range==1,compare_date_data3$n_bydate,NA) %>% 
  na.omit()

t.test(casenum_byday_0,casenum_byday_1)


casenum_byday_0=ifelse(compare_date_data3$time_range==0,compare_date_data3$n_bydate,NA) %>% 
  na.omit()
casenum_byday_1=ifelse(compare_date_data3$time_range==2,compare_date_data3$n_bydate,NA) %>% 
  na.omit()

t.test(casenum_byday_0,casenum_byday_1)


casenum_byday_0=ifelse(compare_date_data3$time_range==1,compare_date_data3$n_bydate,NA) %>% 
  na.omit()
casenum_byday_1=ifelse(compare_date_data3$time_range==2,compare_date_data3$n_bydate,NA) %>% 
  na.omit()

t.test(casenum_byday_0,casenum_byday_1)
```

It's the same as all event circumstances.

### 4. Compare the time range as different seasons and use paired t_test
```{r}
tidydata3 = tidydata3 %>% 
  mutate(
    mon = as.numeric(mon)
  )

compare_date_data4 = tidydata3 %>% 
  filter(level == "FELONY") %>%
  filter(mon > 5 & mon < 9) %>% 
  group_by(year) %>% 
  dplyr::summarize(
    n_bydate = n(),
    covid_state = covid_state
  ) %>% 
  unique()

compare_date_data5 = tidydata3 %>% 
  filter(level == "FELONY") %>%
  filter(mon > 11 | mon < 3) %>% 
  group_by(year) %>% 
  dplyr::summarize(
    n_bydate = n(),
    covid_state = covid_state
  ) %>% 
  unique()

t.test(compare_date_data4$n_bydate,compare_date_data5$n_bydate,paired = TRUE)
```

Here is a significant difference b/w winter and summer.
