---
title: "Statistical Testing"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

\ \par

```{r}
library(tidyverse)
library(purrr)
library(modelr)
library(dbplyr)
library(plyr)
library(lubridate)
```

\ \par

### Data cleaning


```{r}
a <- list.files("./Data")     
a 
dir <- paste("./Data/",a,sep="")                 
n <- length(dir) 
merge_data <- read.csv(file = dir[1],header=T,sep=",")   
merge_data <- cbind(dir[1], merge_data )
merge_data <- rename(merge_data, c("dir[1]"="yyyymm"))
for (i in 2:7){
  new.data <- read.csv(file = dir[i], header=T, sep=",")
  new.data <- cbind(dir[i], new.data)
  new.data <- rename(new.data, c("dir[i]"="yyyymm"))
  merge_data <- rbind(merge_data,new.data)
}
```

### Data cleaning, drop data with unknown
```{r}
# filter data with unknown
tidydata1=merge_data%>%
  na.omit()%>%
  filter(!susp_sex == "U")%>%
  filter(!susp_race == "UNKNOWN")%>%
  filter(!susp_age_group == "UNKNOWN")%>%
   filter(!vic_race == "UNKNOWN")%>%
  filter(!vic_age_group == "UNKNOWN")%>%
  separate(date,into=c("mon","day","year"),sep="/")%>%
  mutate(jurisdiction_code=as.numeric(jurisdiction_code))%>%
  mutate(susp_sex=ifelse(susp_sex=="M",1,0),vic_sex=ifelse(vic_sex=="M",1,0),jurisdiction_code=ifelse(jurisdiction_code>=3,3,jurisdiction_code+0))%>%
  mutate(jurisdiction_code=as.factor(jurisdiction_code))%>%
  mutate(covid_state=ifelse(year>=2020,1,0))
# when year >=2020, we assume the world is in a covid state.
```

```{r}
tidydata3=tidydata1 %>%
  mutate(
   time= substring(time, 1,2)
  )

tidydata3= tidydata3 %>% 
 mutate(
   time1=substring(time,1,1),
   time2=substring(time,2,2),
   time3=ifelse(time1=="0","",time1),
   time4=paste(time3,time2,sep=""),
   time5=as.numeric(time4)
 ) 
  tidydata3=tidydata3 %>% 
  mutate(time_range=case_when(
      (time5 > 22) | (time5 <= 6) ~ 0,
      (time5>6) & (time5<=14) ~ 1,
      (time5>14) & (time5<=22) ~ 2
    ))


```


This is the beginning of test
First of all let's define the "danger" as the _total number of crime event_ occure in _one day_

# dangers define as the criminal events happen by day

## 1.let's see does it correlate to the criminal's gender?
```{r}
crimebyday_men=tidydata1 %>% 
  group_by(year,mon,day) %>% 
  filter(susp_sex==1) %>% 
  dplyr::summarize(
    n_bydate=n(),
  ) %>% 
  unique()

crimebyday_women=tidydata1 %>% 
  group_by(year,mon,day) %>% 
  filter(susp_sex==0) %>% 
  dplyr::summarize(
    n_bydate=n(),
  ) %>% 
  unique()
t.test(crimebyday_men$n_bydate,crimebyday_women$n_bydate)
```
significant difference

## 2.let's see does it correlate to the victim's gender?
```{r}
crimebyday_men=tidydata1 %>% 
  group_by(year,mon,day) %>% 
  filter(vic_sex==1) %>% 
  dplyr::summarize(
    n_bydate=n(),
  ) %>% 
  unique()

crimebyday_women=tidydata1 %>% 
  group_by(year,mon,day) %>% 
  filter(vic_sex==0) %>% 
  dplyr::summarize(
    n_bydate=n(),
  ) %>% 
  unique()
t.test(crimebyday_men$n_bydate,crimebyday_women$n_bydate)
```
significant difference
female victims are significantly higher than male

## 3.let's see does it correlate to the criminals' race?
```{r}
crimebyday_balck=tidydata1 %>% 
  group_by(year,mon,day) %>% 
  filter(susp_race=="BLACK") %>% 
  dplyr::summarize(
    n_bydate=n(),
  ) %>% 
  unique()

crimebyday_white=tidydata1 %>% 
  group_by(year,mon,day) %>% 
  filter(susp_race=="WHITE") %>% 
  dplyr::summarize(
    n_bydate=n(),
  ) %>% 
  unique()
t.test(crimebyday_balck$n_bydate,crimebyday_white$n_bydate)
```
significant
the number of white race criminals are significantly lower than black.

## 4.What about other criminals' race?
```{r}
crimebyday_aspa=tidydata1 %>% 
  group_by(year,mon,day) %>% 
  filter(susp_race=="ASIAN / PACIFIC ISLANDER") %>% 
  dplyr::summarize(
    n_bydate=n(),
  ) %>% 
  unique()

crimebyday_white=tidydata1 %>% 
  group_by(year,mon,day) %>% 
  filter(susp_race=="WHITE") %>% 
  dplyr::summarize(
    n_bydate=n(),
  ) %>% 
  unique()
t.test(crimebyday_aspa$n_bydate,crimebyday_white$n_bydate)

```
The number of ASIAN / PACIFIC ISLANDER race of criminals are less than white criminals.
...if you need more can be shown about race...

## 5.compare crime events number by covid statu
```{r}
crimebyday_covid0=tidydata1 %>% 
  group_by(year,mon,day) %>% 
  dplyr::summarize(
    n_bydate=n(),
    covid_state=covid_state
  ) %>% 
  unique()

casenum_byday_0=ifelse(crimebyday_covid0$covid_state==0,crimebyday_covid0$n_bydate,NA) %>% 
  na.omit()
casenum_byday_1=ifelse(crimebyday_covid0$covid_state==0,NA,crimebyday_covid0$n_bydate) %>% 
  na.omit()
t.test(casenum_byday_0,casenum_byday_1)
```
The diference is significant.
Among the overall status of covid we can see there is a downtrend before covid and after covid.

## 6.compare crime events number by time range
```{r}
crimebyday_time0=tidydata3 %>%
  filter(time_range==0) %>% 
  group_by(year,mon,day) %>% 
  dplyr::summarize(
    n_bydate=n()
  ) %>% 
  unique()

crimebyday_time1=tidydata3 %>%
  filter(time_range==1) %>% 
  group_by(year,mon,day) %>% 
  dplyr::summarize(
    n_bydate=n()
  ) %>% 
  unique()

crimebyday_time2=tidydata3 %>%
  filter(time_range==2) %>% 
  group_by(year,mon,day) %>% 
  dplyr::summarize(
    n_bydate=n()
  ) %>% 
  unique()


t.test(crimebyday_time0$n_bydate,crimebyday_time1$n_bydate)
t.test(crimebyday_time0$n_bydate,crimebyday_time2$n_bydate)
t.test(crimebyday_time1$n_bydate,crimebyday_time2$n_bydate)
```
From the test, we can see a significant different number of criminal events occur at different time range:22-6 6-14 14-22 during a day. And the relation can be conclude as:
22-6 < 6-14 <14-22

# dangers define as the FELONY events happen by day

## about races
```{r}
crimebyday_aspa=tidydata1 %>% 
  filter(level=="FELONY") %>%
  group_by(year,mon,day) %>% 
  filter(susp_race=="ASIAN / PACIFIC ISLANDER") %>% 
  dplyr::summarize(
    n_bydate=n(),
  ) %>% 
  unique()

crimebyday_white=tidydata1 %>% 
  filter(level=="FELONY") %>%
  group_by(year,mon,day) %>% 
  filter(susp_race=="WHITE") %>% 
  dplyr::summarize(
    n_bydate=n(),
  ) %>% 
  unique()
t.test(crimebyday_aspa$n_bydate,crimebyday_white$n_bydate)
```
races is follow the same as normal event assumption

## about covid state
```{r}
compare_date_data3=tidydata3 %>% 
  filter(level=="FELONY") %>% 
  group_by(year,mon,day) %>% 
  dplyr::summarize(
    n_bydate=n(),
    covid_state=covid_state
  ) %>% 
  unique()

casenum_byday_0=ifelse(compare_date_data3$covid_state==0,compare_date_data3$n_bydate,NA) %>% 
  na.omit()
casenum_byday_1=ifelse(compare_date_data3$covid_state==0,NA,compare_date_data3$n_bydate) %>% 
  na.omit()
t.test(casenum_byday_0,casenum_byday_1)
```
significant difference
we can see after covid, the probability of happen FELONY is greater than the time before covid.

## about time range
```{r}
compare_date_data3=tidydata3 %>% 
  filter(level=="FELONY") %>% 
  group_by(year,mon,day,time_range) %>% 
  dplyr::summarize(
    n_bydate=n(),
    covid_state=covid_state
  ) %>% 
  unique()

casenum_byday_0=ifelse(compare_date_data3$time_range==0,compare_date_data3$n_bydate,NA) %>% 
  na.omit()
casenum_byday_1=ifelse(compare_date_data3$time_range==1,compare_date_data3$n_bydate,NA) %>% 
  na.omit()
t.test(casenum_byday_0,casenum_byday_1)

casenum_byday_0=ifelse(compare_date_data3$time_range==0,compare_date_data3$n_bydate,NA) %>% 
  na.omit()
casenum_byday_1=ifelse(compare_date_data3$time_range==2,compare_date_data3$n_bydate,NA) %>% 
  na.omit()
t.test(casenum_byday_0,casenum_byday_1)

casenum_byday_0=ifelse(compare_date_data3$time_range==1,compare_date_data3$n_bydate,NA) %>% 
  na.omit()
casenum_byday_1=ifelse(compare_date_data3$time_range==2,compare_date_data3$n_bydate,NA) %>% 
  na.omit()
t.test(casenum_byday_0,casenum_byday_1)
```
It's the same as all event circumstances.

## Let's compare the time range as different seasons and use paired t_test
```{r}
tidydata3 = tidydata3 %>% 
  mutate(
    mon=as.numeric(mon)
  )

compare_date_data4=tidydata3 %>% 
  filter(level=="FELONY") %>%
  filter(mon>5 & mon<9) %>% 
  group_by(year) %>% 
  dplyr::summarize(
    n_bydate=n(),
    covid_state=covid_state
  ) %>% 
  unique()

compare_date_data5=tidydata3 %>% 
  filter(level=="FELONY") %>%
  filter(mon>11 | mon<3) %>% 
  group_by(year) %>% 
  dplyr::summarize(
    n_bydate=n(),
    covid_state=covid_state
  ) %>% 
  unique()


t.test(compare_date_data4$n_bydate,compare_date_data5$n_bydate,paired=TRUE)

```
Here is a significant difference b/w winter and summer.